---
output: reprex::reprex_document
Title: Forecasting Weighted Averages for Dave
Date: 03/02/2023
knit: reprex::reprex_render
---
1st: we will load libraries.
```{r}
library(readxl)     # Provides functions to read Excel files.
library(dplyr)      # Essential data manipulation functions.
library(lubridate)  # Functions to work with date-times.
library(zoo)        # Contains as.Date() function for irregular time series.
library(forecast)   # Essential time series forecasting tools.
```

2nd: we will import our file.
```{r}
# Define the path to the sales data file.
file <- "your_file_path.csv"

# Check file extension and read in the data accordingly.
ifelse(tools::file_ext(file) == "xlsx", 
       data <- readxl::read_excel(file),
       data <- read.csv(file))
```

Optional, view data.
```{r}
View(data)
```

3rd: we will aggregate sales data by quarter.
```{r}
quarterly_sales <- data %>%
  group_by(
    quarter = quarter(as.Date(paste(YEAR_ID, QTR_ID, "01", sep = "-"))),
    year = year(as.Date(paste(YEAR_ID, QTR_ID, "01", sep = "-")))
  ) %>%
  summarize(sales = sum(SALES))
```

4th: we will create a time series, and a train and validation set.
```{r}
sales.ts <- ts(quarterly_sales$sales,
               start = c(min(quarterly_sales$year), 1),
               frequency = 4)

# Define the validation set as the last 20% of the data.
nValid <- ceiling(0.2 * length(sales.ts))
nTrain <- length(sales.ts) - nValid
train.ts <- window(sales.ts, end = c(tail(quarterly_sales$year, n = 1), nTrain))
valid.ts <- window(sales.ts, start = c(tail(quarterly_sales$year, n = 1), nTrain + 1))
```

5th: we will forecast the time series with different models, and select the best model.
```{r}
# Train multiple time series models.
auto_arima <- auto.arima(train.ts)
tslm <- tslm(train.ts ~ trend + season, lambda = 1)
auto_ets <- ets(train.ts, model = 'ZZZ')

# Forecast using the trained models.
auto_arima_fc <- forecast(auto_arima, h = nValid) 
tslm_fc <- forecast(tslm, h = nValid)
auto_ets_fc <- forecast(auto_ets, h = nValid)

# Evaluate forecast accuracy.
acc_auto_arima <- accuracy(auto_arima_fc, valid.ts) 
acc_tslm <- accuracy(tslm_fc, valid.ts) 
acc_auto_ets <- accuracy(auto_ets_fc, valid.ts) 

# Select the best model based on Root Mean Squared Error (RMSE).
if (acc_auto_arima["Test set", "RMSE"] < acc_tslm["Test set", "RMSE"] && acc_auto_arima["Test set", "RMSE"] < acc_auto_ets["Test set", "RMSE"]) {
  best_model <- auto_arima
} else if (acc_tslm["Test set", "RMSE"] < acc_auto_arima["Test set", "RMSE"] && acc_tslm["Test set", "RMSE"] < acc_auto_ets["Test set", "RMSE"]) {
  best_model <- tslm
} else {
  best_model <- auto_ets
}
```

Optional, check model fit.
```{r}
summary(best_model)
```

6th: Take the best model, and train with full data set.
```{r}
if (class(best_model)[1] == "Arima") {
  model <- auto.arima(sales.ts)
} else if (class(best_model)[1] == "lm") {
  model <- tslm(sales.ts ~ trend + season, lambda = 1)
} else {
  model <- ets(sales.ts, model = 'ZZZ')
}
```

7th: we will make a probability density function with the forecast.
```{r}
fc <- forecast(model)
pdf <- density(fc$mean)
```

Optional, view pdf plot.
```{r}
plot(pdf, ylab = "Probability Density", xlab = "Sales in $", main = "Forecasted PDF of Sales")
```

8th: We will take our pdf and calculate the weighted averages of each region
```{r}
# Define intervals corresponding to different sales regions.
intervals <- list(c(0, 0.1), c(0.1, 0.4), c(0.4, 0.6), c(0.6, 0.9), c(0.9, 1))
weighted_average <- numeric(length(intervals))

# Function to calculate weighted average within a specified interval.
weighted.mean.interval <- function(pdf, interval) {
  prob <- diff(pmax(0, pmin(interval, pdf$x))) / diff(range(pdf$x))
  wtd_avg <- sum(prob * pdf$y[pmax(1, match(interval[1], pdf$x)):pmin(length(pdf$x), match(interval[2], pdf$x))])
  return(wtd_avg)
}

# Compute weighted averages for each interval.
for (i in seq_along(intervals)) {
  weighted_average[i] <- weighted.mean.interval(pdf, intervals[[i]])
}

# View the results.
col_names <- c("0-10%", "10-40%", "40-60%", "60-90%", "90-100%")
weighted_average <- setNames(weighted_average, col_names)
weighted_average
```
